---
title: "Statistics"
output: html_notebook
---

# Mode

```{r}
x <- c(10,10,10,20,20,30,30,40,50)

# table back the frequency of each elements in vector
y <- table(x)

as.numeric(names(y)[which.max(y)])

```

# Range

```{r}
x <- c(7,4,11,8,19)
range(x)

# the difference between Max value & Min value
diff(range(x))
```

# Variance & Standard Deviation

```{r}
x <- c(7,4,11,8,19)
sd(x)

var(x)
```

# MAD (Median Absolute Deviation)

```{r}
x <- c(7,4,11,8,19)
mad(x, constant = 1)
```

# Normal Distribution

### Shapiro test & P-value

```{r}
set.seed(10)
x1 <- rnorm(1000) # 1000 values from standard normal distribution
x2 <- runif(1000) # 1000 values from uniform distribution
x3 <- rgamma(1000,2,1) # 1000 values from gamma distribution
x4 <- c(rep(3,10),1,2,4,5) # Leptokutric distribution

hist(x1, main = 'standard normal distribution')
hist(x2, main = 'uniform distribution')
hist(x3, main = ' gamma distribution')
hist(x4, main = 'Leptokutric distribution')
```

```{r}
# Normal Distribution when Pvalue > 0.05
shapiro.test(x1)$p.value # large p_value follow normal distribution

shapiro.test(x2)$p.value
shapiro.test(x3)$p.value
shapiro.test(x4)$p.value
```

# A- Single Sample

### 1- t_test

-   if your single sample is normally distributed

```{r}
scores <- c(1370,1290,1250,1320,1250,1370,1340,1330,1280,1210,1390,1180,1290,1300,1300,1250,1250,1310,1360)

your.score <- 1330

# to test if normally distributed by Shapiro-wilk test
shapiro.test(scores)
```

```{r}
# t-test because p-value > 0.05
# I don't care about the direction of differency SO "two.sided"
# compare x to mu to get to reject the NUll hypothesis or not
t.test(x=scores, mu=your.score, alternative = "two.sided")

# p_value < 0.05 => SO my_score is significant
```

```{r}
# if we need to get the direction hypothesis // if my_score is higher or lower to my colleges
t.test(x=scores, mu=your.score, alternative = "less")
```

```{r}
t.test(x=scores, mu=your.score, alternative = "greater")
```

-   if the single data not normally distributed

    ```{r}
    score2 <- c(1340,1540,1310,1270,1340,1300,1330,1540,1540,1320)

    your.score2 <- 1300

    shapiro.test(score2)
    ```

-   so we go to Wilicoxon test

    ### 2- Wilicoxon test

    ```{r}
    wilcox.test(score2, mu=your.score2, alternative = "two.sided")
    ```

# B- Two groups

-   1- sample are paired =\> like the same subject but 2 different samples (Healthy & cancer) OR 2 conditions (before and after TTT)

-   2- samples are independent

    ### 1- Paired 2 samples

    -   check the normal distribution on paired difference (group A - group B = paired difference) by Shapiro-wilk normality test

        ```{r}
        set.seed(10)
        before = rnorm(100)
        set.seed(20)
        after = rnorm(100, mean = 2)

        d = after - before

        shapiro.test(d)
        ```

-   p-value of shapiro.test \> 0.05 // so we use t-test

    ```{r}
    t.test(after,before,paired = TRUE, alternative = 'two.sided')
    ```
